Malacology: A Programmable Storage System

### 2016-10-13 paper/paper.tex
paper: add Noah's intro comments
Signed-off-by: Michael Sevilla <mikesevilla3@gmail.com>

\begin{abstract}
Storage systems are caught between rapidly changing data
processing systems and the increasing speed of storage devices. This puts
tremendous pressure on storage systems to adapt both in terms of their
interfaces and their performance. But adapting storage systems can be difficult
because unprincipled changes might jeopardize years of code-hardening and
performance optimization efforts that were necessary for users to entrust their
data to the storage system.  We introduce  Malacology, a prototype programmable
storage system to explore how existing abstractions of common services found in
storage systems can be leveraged to address new data processing systems and the
increasing speed of storage devices. This approach allows a large degree of
flexibility for storage systems to evolve without sacrificing the robustness of
its code-hardened subsystems.  We illustrate the advantages and challenges of
programmability by composing existing primitives into two new higher-level
services: a file system metadata load balancer and a high-performance distributed
shared-log that leverages flash devices. The evaluation demonstrates that our
services inherit desirable qualities of the back-end storage system, including
the ability to make trade-offs for latency/throughput using lease capabilities, balance load, propogate
interfaces, and recover from failure.
\end{abstract}

@@

A number of recent developments are disrupting traditional storage systems.
First, the falling prices of flash storage and the availability of new types of
non-volatile memory that are orders of magnitude faster than traditional
spinning media are moving overall performance bottlenecks away from storage
devices to CPUs and networking, and pressure storage systems to shorten their
code paths and incorporate new
optimizations~\cite{gray_tape_2007,gray_flash_2008}. Second, the demand for
managing structured data and flexible consistency semantics at scale pressure
big data processing systems to use storage abstractions that can meet these
demands~\cite{apache_contributors_parquet_2014}. Finally, production-quality
scalable storage systems available as open source software have established and
are continuing to establish new, \emph{de-facto} API standards at a faster pace
than traditional standards
bodies~\cite{snia_implementing_2014,linux_foundation_kinetic_2015}.

@@

These three trends put evolutionary pressure on storage systems and raise the
question whether there are principles that storage systems designers can follow
to evolve storage systems efficiently and without jeopardizing years of
code-hardening and performance optimization efforts that are important for
users to continue to entrust their data to the storage system.

In this paper we investigate an approach that focuses on generalizing and
composing existing
storage system resources, services, and abstractions that in a generalized form
can be used to \emph{program} new services. By composing higher-level services
over existing abstractions, one can reuse
subsystems and leverage their optimizations, established correctness,
robustness, and efficiency. We will refer to this programmability as
\emph{programmable storage}, which differs from \emph{active storage} (the
injection and execution of arbitrary code in a storage system) and
\emph{software-defined storage} (the control of thin-provisioning of storage).

To illustrate the benefits and challenges of this approach we have designed and
evaluated Malacology, a programmable storage system for constructing new
services by re-purposing existing subsystem abstractions of the storage stack.
We build the framework in Ceph, a popular, open source software storage stack.
We choose Ceph to demonstrate the concept of programmable stoarge because it
offers a broad spectrum of existing services, including distributed locking and
caching services provided by metadata servers, durability and logical storage
devices provided by the backend object store, and propagation of consistent
cluster state provided by the monitoring service (see
Figure~\ref{fig:overview}). As we will show in this paper, this framework is
expressive enough to provide the functionality necessary for implementing new
services. Our contributions are:

{\it \textbf{We define a programmable storage system}} to be a storage system that re-uses and extends existing
abstractions and services of the underlying software stack. Our prototype
includes (but is not limited to) the following building blocks:

@@

\item Versioning interface that uses the underyling storage stack's consensus
infrastructure

\item Logical Device interface for adding functionality to daemons that manage
storage/metadata devices, allowing users to push computation to the data

\item An interface for mediating access to a shared resources, which exposes
serialization primatives using capabilities, caching, and batching

\item Load Balancing interface for managing resources across the cluster in
multi-tenant environments

\item Durability interface for persisting policies using the underyling storage
stack's object store

@@

\item A high-performance distributed shared log service called ZLog which is
based on CORFU~\cite{balakrishnan_corfu_2012}

@@

\section{Highly Tailored, Application-Specific Storage Stacks
}\label{highly-tailored-and-application-specifc-storage-systems}

>>>>

\begin{abstract}
Storage systems are caught between rapidly changing data
processing systems and the increasing speed of storage devices. This puts
tremendous pressure on storage systems to adapt both in terms of their
interfaces and their performance. But adapting storage systems can be difficult
because unprincipled changes might jeopardize years of code-hardening and
performance optimization efforts that were necessary for users to entrust their
data to the storage system.  We introduce  Malacology, a prototype programmable
storage system to explore how existing abstractions of common services found in
storage systems can be leveraged to address new data processing systems and the
increasing speed of storage devices. This approach allows a large degree of
flexibility for storage systems to evolve without sacrificing the robustness of
their code-hardened subsystems.  We illustrate the advantages and challenges of
programmability by composing existing primitives into two new higher-level
services: a file system metadata load balancer and a high-performance distributed
shared-log that leverages flash devices. The evaluation demonstrates that our
services inherit desirable qualities of the back-end storage system, including
the ability to balance load, efficiently propogate cluster metadata, recover
from failure, and to make trade-offs for latency/throughput using leases.
\end{abstract}

@@

A number of recent developments have disrupted traditional storage systems.
First, the falling prices of flash storage and the availability of new types of
non-volatile memory that are orders of magnitude faster than traditional
spinning media are moving overall performance bottlenecks away from storage
devices to CPUs and networking, and pressure storage systems to shorten their
code paths and incorporate new
optimizations~\cite{gray_tape_2007,gray_flash_2008}. Second, the demand for
managing structured data and flexible consistency semantics at scale pressure
big data processing systems to use storage abstractions that can meet these
demands~\cite{apache_contributors_parquet_2014}. Finally, production-quality
scalable storage systems available as open source software have established and
are continuing to establish new, \emph{de-facto} API standards at a faster pace
than traditional standards
bodies~\cite{snia_implementing_2014,linux_foundation_kinetic_2015}.

@@

These three trends put evolutionary pressure on storage systems and raise the
question of whether there are principles that storage systems designers can follow
to evolve storage systems efficiently and without jeopardizing years of
code-hardening and performance optimization efforts that are important for
users to continue to entrust their data to the storage system.

In this paper we investigate an approach that focuses on generalizing and
composing existing
storage system resources, services, and abstractions that in a generalized form
can be used to \emph{program} new services. By composing higher-level services
over existing abstraction one can reuse
subsystems and leverage their optimizations, established correctness,
robustness, and efficiency. We will refer to this programmability as
\emph{programmable storage}, which differs from \emph{active storage} (the
injection and execution of arbitrary code in a storage system) and
\emph{software-defined storage} (the control of thin-provisioning of storage).

To illustrate the benefits and challenges of this approach we have designed and
evaluated Malacology, a programmable storage system for constructing new
services by re-purposing existing subsystem abstractions of the storage stack.
We build the Malacology in Ceph, a popular open source software storage stack.
We choose Ceph to demonstrate the concept of programmable storage because it
offers a broad spectrum of existing services, including distributed locking and
caching services provided by metadata servers, durability and object interfaces
provided by the backend object store, and propagation of consistent
cluster state provided by the monitoring service (see
Figure~\ref{fig:overview}). As we will show in this paper, Malacology is
expressive enough to provide the functionality necessary for implementing new
services. Our contributions are:

{\it \textbf{The definition a programmable storage system}} to be "a storage
system that allows exisiting abstractions to be re-used and composed to form
new data and metadata management services whose operation is orothogonal to
system correctness".  Our prototype includes (but is not limited to) the
following building blocks:

@@

\item An interface for managing strongly-consistent time-varying service
metadata.

\item An interface for installing and evolving domain-specific, cluster-wide data
interfaces.

\item An interface for managing access to shared resources using a variety of
optimization strategies.

\item An interface for load balancing resources across the cluster.

\item A durability interface for persisting policies using the underyling storage
stack's object store

@@

\item A high-performance distributed shared log service called ZLog, that is an
implementation of CORFU~\cite{balakrishnan_corfu_2012}

@@

\section{Application-Specific Storage Stacks
}\label{highly-tailored-and-application-specifc-storage-systems}
